---
title: "Projections for Following Semesters "
date: "`r format(Sys.time(), '%d %B, %Y')`"
---


## Years

Years chosen: `r input$years[1]` to `r input$years[2]`.

## Projection Target

Target variable chosen: `r input$target`.

## Forcast Model

Forecast model chosen: `r input$model`.

## Projections Graph
```{r plot, echo=FALSE, fig.width=8, fig.height=6}
#---------------------------------------------------------#
# the following code is the same as renderPlot in server.R
# Update it when updating server.R, just copy and paste.
#---------------------------------------------------------#


# obtain desired and preed data
desired_dat = grabData()$desiredData
pred_dat = grabData()$predData

cols = c("0" = "blue", "1" = "red")

# choose prediction variable

if (input$target == "Enrollment"){
    pred_var = "enrollment"
    lo = 1000
    hi = 6000
} else if (input$target == "Attempted Credits") {
    pred_var = "credits"
    lo = 10000
    hi = 65000
} else{
    pred_var = "hours"
    lo = 400000
    hi = 1350000
}


if (input$model == "Basic Linear Regression"){
    formula = paste0(pred_var, " ~ year + unemployment_rate_last_year + semesterFall")
    
    fit = lm(formula, data = desired_dat)
    
    # make predictions on desired data to see the performance
    desired_dat$prejection_LR = fit$fitted.values
    
    # make predictions for the following semesters
    pred_dat$prejection_LR = predict.lm(fit, newdata = pred_dat) 
    
    # merger two data sets
    merge_dat = rbind(desired_dat %>% select(year, semesterFall, unemployment_rate_last_year, prejection_LR), 
                      pred_dat)
    
    # making plots again
    plt = ggplot(aes(x = year, y = prejection_LR), data = merge_dat) +
        geom_point(aes(shape = as.factor(semesterFall))) + 
        scale_x_continuous(breaks = seq(min(merge_dat$year), max(merge_dat$year), 1),
                           labels = seq(min(merge_dat$year), max(merge_dat$year), 1)) + 
        scale_shape_manual(name = "Semester\n Projections",
                           values = c("0" = 0, "1" = 2),
                           breaks = c("0", "1"),
                           labels = c("Spring", "Fall")) + 
        geom_point(aes(x = year, y = get(pred_var), color = as.factor(semesterFall)), size = 2, data = desired_dat) + 
        scale_color_manual(name = "Semester",
                           values = c("0" = "red", "1" = "blue"),
                           breaks = c("0", "1"),
                           labels = c("Spring", "Fall")) +
        xlab("Year") + 
        ylab(input$target) + 
        ylim(c(lo, hi)) + 
        myTheme
    print(plt)
} else if(input$model == "Prior Moving Average (3yrs)"){
    # average 3 previous projection data
    # projection separatedly
    shift = 3
    
    # projection for Spring
    actual_spring = desired_dat %>% filter(semesterFall == 0) %>% 
        select(year, semesterFall, contains(pred_var))
    pred_spring = pred_dat %>% filter(semesterFall == 0) %>%
        select(year, semesterFall)
                
    actual_s = pma(actual_spring, pred_spring, pred_var, shift)[[1]] 
    pred_s = pma(actual_spring, pred_spring, pred_var, shift)[[2]] # only return 3 variables
    
    # projection for Fall 
    actual_fall = desired_dat %>% filter(semesterFall == 1) %>% 
        select(year, semesterFall, contains(pred_var))
    pred_fall = pred_dat %>% filter(semesterFall == 1) %>%
        select(year, semesterFall)
    
    actual_f = pma(actual_fall, pred_fall, pred_var, shift)[[1]] 
    pred_f = pma(actual_fall, pred_fall, pred_var, shift)[[2]] # only return 3 variables
    
    
    actual = rbind(actual_s, actual_f)
    preds = rbind(pred_s, pred_f)
    
    # merger two data sets
    merge_dat = rbind(actual, preds)
    
    # making plots again
    plt = ggplot(aes(x = year, y = projection_PMA_3yr), data = merge_dat) +
        geom_point(aes(shape = as.factor(semesterFall))) + 
        scale_x_continuous(breaks = seq(min(merge_dat$year), max(merge_dat$year), 1),
                           labels = seq(min(merge_dat$year), max(merge_dat$year), 1)) + 
        scale_shape_manual(name = "Semester\n Projections",
                           values = c("0" = 0, "1" = 2),
                           breaks = c("0", "1"),
                           labels = c("Spring", "Fall")) + 
        geom_point(aes(x = year, y = get(pred_var), color = as.factor(semesterFall)), size = 2, data = desired_dat) + 
        scale_color_manual(name = "Semester",
                           values = c("0" = "red", "1" = "blue"),
                           breaks = c("0", "1"),
                           labels = c("Spring", "Fall")) +
        xlab("Year") + 
        ylab(input$target) + 
        ylim(c(lo, hi)) + 
        myTheme
    print(plt)
    
} else if(input$model == "Prior Moving Average (5yrs)"){
    # average 5 previous projection data
    shift = 5
    
    # projection for Spring
    actual_spring = desired_dat %>% filter(semesterFall == 0) %>% 
        select(year, semesterFall, contains(pred_var))
    pred_spring = pred_dat %>% filter(semesterFall == 0) %>%
        select(year, semesterFall)
    
    actual_s = pma(actual_spring, pred_spring, pred_var, shift)[[1]] 
    pred_s = pma(actual_spring, pred_spring, pred_var, shift)[[2]] # only return 3 variables
    
    # projection for Fall 
    actual_fall = desired_dat %>% filter(semesterFall == 1) %>% 
        select(year, semesterFall, contains(pred_var))
    pred_fall = pred_dat %>% filter(semesterFall == 1) %>%
        select(year, semesterFall)
    
    actual_f = pma(actual_fall, pred_fall, pred_var, shift)[[1]] 
    pred_f = pma(actual_fall, pred_fall, pred_var, shift)[[2]] # only return 3 variables
    
    
    actual = rbind(actual_s, actual_f)
    preds = rbind(pred_s, pred_f)
    
    # merger two data sets
    merge_dat = rbind(actual, preds)
    
    # making plots again
    plt = ggplot(aes(x = year, y = projection_PMA_5yr), data = merge_dat) +
        geom_point(aes(shape = as.factor(semesterFall))) + 
        scale_x_continuous(breaks = seq(min(merge_dat$year), max(merge_dat$year), 1),
                           labels = seq(min(merge_dat$year), max(merge_dat$year), 1)) + 
        scale_shape_manual(name = "Semester\n Projections",
                           values = c("0" = 0, "1" = 2),
                           breaks = c("0", "1"),
                           labels = c("Spring", "Fall")) + 
        geom_point(aes(x = year, y = get(pred_var), color = as.factor(semesterFall)), size = 2, data = desired_dat) + 
        scale_color_manual(name = "Semester",
                           values = c("0" = "red", "1" = "blue"),
                           breaks = c("0", "1"),
                           labels = c("Spring", "Fall")) +
        xlab("Year") + 
        ylab(input$target) + 
        ylim(c(lo, hi)) + 
        myTheme
    print(plt)
}
```


## Projections Output

```{r fun, echo=FALSE}
semester_fun <- function(x){
    if (x == 1){
        return("Fall")
    } else {
        return("Spring")
    }
}

```


```{r model, echo=FALSE, collapse=TRUE}
#--------------------------------------------------------------#
# the following code is the same as renderDataTable in server.R
# Update it when updating server.R, just copy and paste.
#--------------------------------------------------------------#

# obtain desired and preed data
desired_dat = grabData()$desiredData
pred_dat = grabData()$predData

# choose prediction variable

if (input$target == "Enrollment"){
    pred_var = "enrollment"
    lo = 1000
    hi = 6000
} else if (input$target == "Attempted Credits") {
    pred_var = "credits"
    lo = 10000
    hi = 65000
} else{
    pred_var = "hours"
    lo = 500000
    hi = 1500000
}

if (input$model == "Basic Linear Regression"){
    formula = paste0(pred_var, " ~ year + unemployment_rate_last_year + semesterFall")
    
    fit = lm(formula, data = desired_dat)
    
    # make predictions on desired data to see the performance
    desired_dat$prejection_LR = fit$fitted.values
    
    # make predictions for the following semesters
    pred_dat$prejection_LR = predict.lm(fit, newdata = pred_dat) 
    
    # merger two data sets
    merge_dat = rbind(desired_dat %>% select(year, semesterFall, unemployment_rate_last_year, prejection_LR), 
                      pred_dat)
    
    # print the projections
    print(pred_dat %>% select(-unemployment_rate_last_year))
    
} else if(input$model == "Prior Moving Average (3yrs)"){
    # average 3 previous projection data
    shift = 3
    
    # projection for Spring
    actual_spring = desired_dat %>% filter(semesterFall == 0) %>% 
        select(year, semesterFall, contains(pred_var))
    pred_spring = pred_dat %>% filter(semesterFall == 0) %>%
        select(year, semesterFall)
    
    actual_s = pma(actual_spring, pred_spring, pred_var, shift)[[1]] 
    pred_s = pma(actual_spring, pred_spring, pred_var, shift)[[2]] # only return 3 variables
    
    # projection for Fall 
    actual_fall = desired_dat %>% filter(semesterFall == 1) %>% 
        select(year, semesterFall, contains(pred_var))
    pred_fall = pred_dat %>% filter(semesterFall == 1) %>%
        select(year, semesterFall)
    
    actual_f = pma(actual_fall, pred_fall, pred_var, shift)[[1]] 
    pred_f = pma(actual_fall, pred_fall, pred_var, shift)[[2]] # only return 3 variables
    
    
    actual = rbind(actual_s, actual_f)
    preds = rbind(pred_s, pred_f)
    print(preds)
    
} else if(input$model == "Prior Moving Average (5yrs)"){
    # average 5 previous projection data
    shift = 5
    
    # projection for Spring
    actual_spring = desired_dat %>% filter(semesterFall == 0) %>% 
        select(year, semesterFall, contains(pred_var))
    pred_spring = pred_dat %>% filter(semesterFall == 0) %>%
        select(year, semesterFall)
    
    actual_s = pma(actual_spring, pred_spring, pred_var, shift)[[1]] 
    pred_s = pma(actual_spring, pred_spring, pred_var, shift)[[2]] # only return 3 variables
    
    # projection for Fall 
    actual_fall = desired_dat %>% filter(semesterFall == 1) %>% 
        select(year, semesterFall, contains(pred_var))
    pred_fall = pred_dat %>% filter(semesterFall == 1) %>%
        select(year, semesterFall)
    
    actual_f = pma(actual_fall, pred_fall, pred_var, shift)[[1]] 
    pred_f = pma(actual_fall, pred_fall, pred_var, shift)[[2]] # only return 3 variables
    
    
    actual = rbind(actual_s, actual_f)
    preds = rbind(pred_s, pred_f)
    print(preds)
}
```

The projection of `r input$target` for `r pred_dat[1, 'year']` `r semester_fun(pred_dat[1, 'semesterFall'])` is `r round(pred_dat[1, names(pred_dat) %in% c("projection_PMA_5yr", "projection_PMA_3yr", "prejection_LR")],0)`. 

Note: variable _semesterFall_ is an indicator for Fall semester. 1 indicates for Fall semester, 0 for Spring semester.





